---
title: "p8105_hw2_xj2249"
author: "jx"
date: "2019/9/24"
output: github_document
---

# Problem 1
## Read and clean the Mr. Trash Wheel sheet
```{r, message = FALSE}
library(tidyverse)
mr_trash <- readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                               sheet = "Mr. Trash Wheel",
                               range = "A2:N336") %>% 
        janitor::clean_names() %>% 
        drop_na(date) %>% 
        mutate(sports_balls = as.integer(sports_balls))
```

## Read and clean precipitation data for 2017 and 2018
```{r, message = FALSE}
pre_18 <- readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                             sheet = "2018 Precipitation",
                             range = "A2:B14") %>% 
        janitor::clean_names() %>% 
        drop_na(total) %>% 
        mutate(year = "2018")
pre_17 <- readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                             sheet = "2017 Precipitation",
                             range = "A2:B14") %>% 
        janitor::clean_names() %>% 
        drop_na(total) %>% 
        mutate(year = "2017")
pre_17_18 <- rbind(pre_18,pre_17) %>% 
        mutate(month = month.name[month])
```


Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. 

For available data, what was the total precipitation in 2018? 


What was the median number of sports balls in a dumpster in 2017?  done 

The median number of sports balls in a dumpster in 2017 is  `r median(pull(filter(mr_trash,year == 2017),sports_balls))`


# Problem 2
## First, clean the data in pols-month.csv
```{r, message = FALSE}
pols <- 
        read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
        separate(mon,into = c("year","month","day"), sep = "-", convert = TRUE) %>% 
        mutate(month = month.name[month],
               president = ifelse(prez_dem == 1,"dem","gop")) %>%
        select(-c("prez_dem","prez_gop","day"))

```

## Second, clean the data in snp.csv 
```{r, message = FALSE}
snp <- 
        read_csv("./data/fivethirtyeight_datasets/snp.csv") %>%
        separate(date,into = c("month","day","year"), sep = "/", convert = TRUE) %>% 
        select("year","month","close") %>% 
        arrange(year,month) %>% 
        mutate(month = month.name[month])
        
```

## Third, tidy the unemployment data 
This process will involve switching from ???wide??? to ???long??? format; ensuring that key variables have the same name; and ensuring that key variables take the same values.
```{r, message = FALSE}
unemployment <- 
        read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
        janitor::clean_names() %>% 
        pivot_longer("jan":"dec", names_to = "month",values_to = "umemployment_%") %>% 
        mutate(month = month.name[match(month,tolower(month.abb))])
```

## Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r}
pol_snp <- left_join(pols,snp,by = c("year", "month"))
pol_snp_unemp <- left_join(pol_snp,unemployment,by = c("year", "month"))
```


# Problem 3
## Load and tidy the data
```{r}
baby_name <- 
        read_csv("./data/Popular_Baby_Names.csv") %>% 
        janitor::clean_names() %>% 
        mutate(childs_first_name = str_to_title(childs_first_name),
               ethnicity = str_to_sentence(recode(ethnicity, "ASIAN AND PACIFIC ISLANDER" = "ASIAN AND PACI",
                                                                "BLACK NON HISPANIC" = "BLACK NON HISP",
                                                                        "WHITE NON HISPANIC" = "WHITE NON HISP"))) %>% 
        distinct(.keep_all = TRUE) %>% 
        view()

```


## Creat table for "¡°Olivia¡±
```{r, message = FALSE}
library(knitr)
pop_olivia <- 
        baby_name %>% 
        filter(childs_first_name == "Olivia") %>% 
        select("year_of_birth","ethnicity","rank") %>% 
        pivot_wider(names_from = year_of_birth, values_from = rank ) %>% 
        select("ethnicity","2011","2012","2013","2014","2015","2016")

kable(pop_olivia) 


# Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).


```

## The most popular name among male children over time.
```{r}
pop_male_name <- 
        baby_name %>% 
        filter(gender == "MALE" & rank == 1) %>% 
        select(year_of_birth, ethnicity, childs_first_name) %>% 
        pivot_wider(names_from = year_of_birth, values_from = childs_first_name ) %>% 
        select("ethnicity","2011","2012","2013","2014","2015","2016") 
kable(pop_male_name)
```

## A scatter plot 
```{r}
baby_name %>%
        filter(gender == "MALE" & year_of_birth == 2016 & ethnicity == "White non hisp") %>% 
        ggplot(aes(x = rank, y = count)) + 
        geom_text(aes(label = childs_first_name),hjust = 0, vjust = 0,size = 3) +
        geom_point(position = "jitter",alpha = 0.25, size = 1) 

```

