---
title: "p8105_hw2_xj2249"
author: "jx2249"  

date: "2019/9/24"
output: github_document
---

# Problem 1
## Read and clean the Mr. Trash Wheel sheet
```{r, message = FALSE}
library(tidyverse)
mr_trash <- 
        readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                               sheet = "Mr. Trash Wheel",
                               range = "A2:N336") %>% 
        janitor::clean_names() %>% 
        drop_na(dumpster) %>% 
        mutate(sports_balls = as.integer(sports_balls))
```

## Read and clean precipitation data for 2017 and 2018
```{r, message = FALSE}
pct_18 <- 
        readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                             sheet = "2018 Precipitation",
                             range = "A2:B14") %>% 
        janitor::clean_names() %>% 
        drop_na(total) %>% 
        mutate(year = "2018")

pct_17 <- 
        readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                             sheet = "2017 Precipitation",
                             range = "A2:B14") %>% 
        janitor::clean_names() %>% 
        drop_na(total) %>% 
        mutate(year = "2017")

pct_17_18 <- 
        rbind(pct_18,pct_17) %>% 
        mutate(month = month.name[month])
```


```{r, include = FALSE }
skimr::skim(mr_trash)
```
* In "Mr. Trash Wheel" dataset, there are 285 observations, embodying information on the dumpter number, date of collection, amount of total litter and litter type.
* It contains information of 285 dumpsters dating from 2014 to 2018. The average litter weight and volume are 3.33 tons/year and 15 cubic yards/year.
* And the median number of sports balls in a dumpster in 2017 is  `r median(pull(filter(mr_trash,year == 2017),sports_balls))`.
* In "precipitation(2017-2018)" dataset,  there are `r nrow(pct_17_18)` observations, with 3 variables, showing the total precipitation data for 19 months in 2017 and 2018. 
* The total precipitation in 2017 and 2018 are `r sum(pull(filter(pct_17_18,year == 2017),total))`
and `r sum(pull(filter(pct_17_18,year == 2018),total))` respectively. Besides, the average precipitation in 2017 and 2018 are `r round(mean(pull(filter(pct_17_18,year == 2017),total)), digits=2) ` and `r round(mean(pull(filter(pct_17_18,year == 2018),total)), digits=2)` respectively.


# Problem 2
## First, clean the data in pols-month.csv
```{r, message = FALSE}
pols <- 
        read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
        separate(mon,into = c("year","month","day"), sep = "-", convert = TRUE) %>% 
        mutate(month = month.name[month],
               president = ifelse(prez_dem == 1,"dem","gop")) %>%
        select(-c("prez_dem","prez_gop","day"))

```

## Second, clean the data in snp.csv 
```{r, message = FALSE}
snp <- 
        read_csv("./data/fivethirtyeight_datasets/snp.csv") %>%
        separate(date,into = c("month","day","year"), sep = "/", convert = TRUE) %>% 
        select("year","month","close") %>% 
        arrange(year,month) %>% 
        mutate(month = month.name[month])
        
```

## Third, tidy the unemployment data.
```{r, message = FALSE}
unemployment <- 
        read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
        janitor::clean_names() %>% 
        pivot_longer("jan":"dec", names_to = "month",values_to = "umemployment_%") %>% 
        mutate(month = month.name[match(month,tolower(month.abb))])
```

## Finally, join the datasets.
```{r}
pol_snp <- left_join(pols,snp,by = c("year", "month"))
pol_snp_unemp <- left_join(pol_snp,unemployment,by = c("year", "month"))
```

* The "pols-month" dataset contains 9 variables concerning whether the president
was republican or democratic and the number of governors,senators and representatives
who are democratic or republican from Jan 1974 to June 2015.
 
* In "snp" dataset, the closing value of the S&P stock index and its associated date are both included. 

* The "unemployment" dataset records the percentage of unemployment in each month from 1948 to 2015.

```{r, include = FALSE }
skimr::skim(pol_snp_unemp)
```

* The resulting dataset has a 822 x 11 dimension, namely 822 obervations and 11 variables. 
It covers data on some key variables, including "president","gov_dem","rep_dem","rep_gop","sen_dem","sen_gop","close" and "umemployment_%"** from year of **1947** to **2015**. 


# Problem 3
## Load and tidy the data
```{r}

baby_name <- 
        read_csv("./data/Popular_Baby_Names.csv") %>% 
        janitor::clean_names() %>% 
        mutate(childs_first_name = str_to_title(childs_first_name),
               ethnicity = str_to_sentence(recode(ethnicity, "ASIAN AND PACIFIC ISLANDER" = "ASIAN AND PACI",
                                                                "BLACK NON HISPANIC" = "BLACK NON HISP",
                                                                        "WHITE NON HISPANIC" = "WHITE NON HISP")),
               gender = str_to_title(gender)
               ) %>% 
        distinct(.keep_all = TRUE) %>% 
        view()
```


## Creat table for "Olivia"
```{r, message = FALSE}
pop_olivia <- 
        baby_name %>% 
        filter(childs_first_name == "Olivia") %>% 
        select("year_of_birth","ethnicity","rank") %>% 
        pivot_wider(names_from = year_of_birth, values_from = rank ) %>% 
        select("ethnicity","2011","2012","2013","2014","2015","2016")
knitr::kable(pop_olivia) 
```


## The most popular name among male children over time.
```{r}
pop_male_name <- 
        baby_name %>% 
        filter(gender == "Male" & rank == 1) %>% 
        select(year_of_birth, ethnicity, childs_first_name) %>% 
        pivot_wider(names_from = year_of_birth, values_from = childs_first_name ) %>% 
        select("ethnicity","2011","2012","2013","2014","2015","2016") 
knitr::kable(pop_male_name) 
```

## Scatter plot--rank and count of male names in white non-hispanic children
```{r, fig.height = 6, fig.width=6}
baby_name %>%
        filter(gender == "Male" & year_of_birth == 2016 & ethnicity == "White non hisp") %>% 
        ggplot(aes(x = rank, y = count)) + 
        geom_point(position = "jitter",alpha = 0.25, size = 1) +
        geom_text(aes(label = childs_first_name),hjust = 0, vjust = 0,size = 3) +
        labs(title = "Rank and count of male names in white non-hispanic children") +
        theme(plot.title = element_text(hjust = 0.5))
```

